Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-11-01 GWLDCII.tex
all small comments on M addressed

We summarize here only the subset of positroid theory that we require in this paper; the interested reader is referred to \cite{Postnikov} for more details.

@@

For our purposes, a {\bf positroid} is\footnote{Note that matroid isomorphism allows arbitrary permutations of the ground set but positroid isomorphism allows only cyclic permutations, so the property of being a positroid is not in general preserved by matroid isomorphism.} a matroid $M$ (with ground set $[n]$ and bases $\cB \subseteq \binom{[n]}{k}$) which can be represented by an element of the positive Grassmannian $\Gr(k,n)$. In other words, there exists a full-rank $k\times n$ real matrix whose $k\times k$ minors are all nonnegative and such that the minor given by the columns of $J$ 
is positive if and only if $J \in \cB$.

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces and Le diagrams, which we introduce below. In order to do so, we first need some preliminary definitions.

@@

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} The independent sets of the matroid $M(W)$ associated to an admissible Wilson loop diagram $W = (\cP,[n])$ and realized by $C(W)$ are exactly those subsets $V \subseteq [n]$ such that $\nexists U\subseteq V$ satisfying $|\Prop(U)| < |U|$. \end{thm}
In other words, the independent sets of $M(W)$ correspond to sets of vertices in $W$ in which no subset supports fewer propagators than the vertices it contains.

@@

  Because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  Consequently, a choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statement, hence proving the result.
\end{proof}

@@

By relating the behavior of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Marin-Amat also showed that $M(W)$ is in fact a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}.

However, to go from the Wilson loop diagram $W$ to the positroid cell, one had to go through the matroid explicitly realized via $C(W)$, then make a list of all its nonzero $k\times k$ minors and extract the Le diagram or Grassmann necklace via the methods described in Section~\ref{sec:positroid background}. In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible Wilson loop diagram to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

\begin{dfn}\label{props inside p}
Let $W = (\cP, [n])$ be a weakly admissible Wilson loop diagram, with $p \in \cP$ supported on edges $i$ and $j$. Write $(p, <_i)$ to represent the same propagator directed from $i$ to $j$. Then the sets of propagators {\bf inside} and {\bf outside} of $(p,<_i)$ are defined to be
\begin{gather*}\cP_{in}(p,<_i) = \{ (k,l) \in \cP \ |\ i \leq_i k <_i l \leq_i j \}, \\
\cP_{out}(p,<_i) = \cP\setminus \cP_{in}(p,<_i),
\end{gather*}
respectively.
\end{dfn}
Note that we have $p\in \cP_{in}(p, <_i)$ and so $\cP_{in}(p, <_i) = \cP_{out}(p, <_j)\cup\{p\}$.

@@

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible Wilson loop diagram with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two things occurs:
  \vspace{-0.5em}
  \begin{enumerate}
    \item $W$ has a propagator of length $\leq 6$ with a propagator of length 2 on one side of it and nothing else on that side.\label{item big and 2}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator ends between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
See Figure~\ref{fig: small configs} for an illustration of the two configurations.
\end{lem}

@@

We will inductively construct a sequence of propagators $p_r$ and $q_r$, with $\ell(p_r) = 2$ for each $r$ and such that either of the following are true:
\begin{itemize}
\item $p_r$ forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement, at which point the induction terminates, 
  %a configuration \eqref{eq:2 prop config} in $D$
{\em or} 
\item there is a propagator $q_r$ satisfying $\ell(q_r) \geq 4$, and $\{p_1, \ldots, p_r, q_1, \ldots, q_{r-1}\}$ are all on the same side of $q_r$.  
\end{itemize}

@@

The first case is that $i_r \leq j_{r+1} \leq i_r+2$ and $k_r-2 \leq j_{r+1}+2 \leq k_{r}$ (so $p_{r+1}$ has one end before $i_{r} + 3$ and the other after $k_{r} - 2$, i.e. there are at most two vertices between an endpoint of $p_{r+1}$ and an endpoint of $q_r$).  Then since $p_{r+1}$ has length 2, it must be that $(i_r+3)+1 \geq k_r-2$ and so $q_r$ has length $\leq 6$.  By the minimality assumption on $q_r$, no propagator in $W_r$ has $q_r$ on one side and $p_{r+1}$ on the other side. Therefore, any propagator in $W_r$ other than $q_r$ and $p_{r+1}$ must be of the form $(i,j)$ with $i_r\leq i < j\leq i_r+2$ or $k_r-2 \leq i < j\leq k_r$. Both of these require that $(i,j)$ has length 2, and so we have configuration \ref{item pair of 2s} which we have already assumed does not occur.  Consequently, $W_r$ contains only the two propagators $q_r$ and $p_{r+1}$, yielding configuration \ref{item big and 2} from the lemma statement and again contradicting our assumption. Therefore this first case cannot occur.

The second case is that there are at least $3$ vertices between one endpoint of $p_{r+1}$ and the nearest endpoint of $q_r$. In other words, at least one of the following four things happen:
\begin{itemize} 
\item $i_r+3 \leq j_{r+1}\leq k_r-3$ (one endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest endpoint of $q_r$), or
\item $i_r+3 \leq j_{r+1}+2\leq k_r-3$ (the other endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest endpoint of $q_r$), or
\item $i_r = j_{r+1}$ ($p_{r+1}$ and $q_r$ both have an end on edge $i_r$), or 
\item $k_r = j_{r+1} +2$  ($p_{r+1}$ and $q_r$ both have an end on edge $k_r$).
\end{itemize}
In other words, at least one end of $p_{r+1}$ is supported entirely by vertices in the interval $[i_r+3,k_r+2]$ (the first two bullet points) or both ends lie in $[i_r, i_r+3]$ (third point) or both ends lie in $[k_r-2, k_r+1]$ (fourth point). These situations all behave similarly, and by symmetry it suffices to only consider the second and third possibilities.

We have $j_{r+1}+4 \leq k_r-1$ in both of these cases: in the situation of the second bullet point this follows directly from the inequality, while in the third case it follows from the fact that $\ell(p_{r+1}) = 2$ and $\ell(q_r) \geq 4$. Therefore the vertex $j_{r+1}+4$ is in the interval $[i_r, k_r+1]$ but does not support either $q_r$ or $p_{r+1}$, and must therefore support some other propagator $t$.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and $W_r$ has no propagators of length 3, it follows that $\ell(t) \geq 4$.  If $q_r$ and $p_{r+1}$ were on different sides of $t$ then this would contradict the minimality of $q_r$.  Therefore we can orient $t$ so that all previous $q_i$ and $p_i$ belong to $\cP_{out}(t)$, and set $q_{r+1} = t$ to continue the induction.

The overall result then follows by induction.
\end{proof}

@@

In this section, we give an algorithm for passing directly from the Wilson loop diagrams to its Grassmann necklace. This not only greatly simplifies the previously known process to obtain a positroid cell from a Wilson loop diagram, but will also allow us to relate the behavior of the positroid $M(W)$ directly to the configuration of propagators in $W$. In Section~\ref{sec poles} we use these results to show that the denominator $R(W)$ of the integrand associated to a Wilson loop diagram can be written in terms of the Grassmann necklace of a diagram.

@@

\begin{dfn} \label{proporderingdfn}
  There is an ordering on the propagators on $W = (\cP, [n])$ defined by $I_i$. This is the ordering in which they contribute to the algorithm as it constructs $I_i$.
%  $\cP = \{p_1, \ldots p_k\}$, where 
% \[I_i = \{I_i(p_1) <_i I_i(p_2) <_i \ldots <_i I_i(p_k)\}.\]
% (The superscript $W$ is omitted here to simplify the notation.)\note{Sian tweaked this slightly to emphasise the $<_i$ order, because standard convention in most papers is to write all GN terms in the 1-order.}
\end{dfn}

@@

\begin{lem}\label{lem no fourth vertex}
Let $W$ be an admissible Wilson loop diagram containing at least one propagator. For any $i \in [n]$ and for any $p \in \cP$, $I_i(p)$ is not maximal (with respect to $<_i$) amongst the vertices that support $p$.
\end{lem} 
\begin{proof}
Suppose for contradiction that we have $i \in [n]$ and $p \in \cP$ such that $I_i(p)$ is $<_i$-maximal in $V(p)$. There are two possibilities, which are illustrated in Figure~\ref{fig:no fourth vertex}: either $p = (i-1,b)$ for some $b$ (so $i$ and $I_i(p) = i-1$ are the two ends of a single edge), or $p = (a,b)$ with $i\leq_i a <_i b$ and $I_i(p) = b+1$.

In Case 1, observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $I_i$ would assign $p$ to $b+1$. Now $q$ also contributes the $<_i$-maximal vertex in its support to $I_i$ and is an instance of Case 2. Therefore it suffices to show that no admissible Wilson loop diagram admits the configuration illustrated in Case 2 of Figure~\ref{fig:no fourth vertex}.

@@

In order for $q$ to remain unassigned until vertex $b$, there must be another propagator $r$ with an end on edge $d$ and $I_i(r) = d+1$; the only way this can occur is if $r$ is outside of $(q,<_i)$ but inside $(p,<_i)$. But now $r$ contributes its fourth vertex to $I_i$, again contradicting the minimality of $p$.
\end{proof}

@@

\begin{rmk}\label{rmk cyclic}
  Before the proof, let us observe that one of the main ways that the above lemma will be used is in observations like the following.  Suppose $p=(i,j)$ is a propagator and $I_{j+1}(p) = j+1$.  Consider $I_{j+2}(p)$.  Since $j+1$ is the last vertex in the $j+2$ order, $p$ can not contribute $j+1$ to $I_{j+2}$, so the fact that the intervals in the statement of the lemma are non-empty and occur in their cyclic order implies that $I_{j+2}(p)=i$.  More generally, we know that whenever a propagator stops contributing a particular vertex, its next contribution must be its cyclically next vertex.  This little observation is surprisingly useful, to the point that the details which appear in the proof of Lemma~\ref{vertex cyclic int lem} are never needed again: the mere fact of non-empty intervals in the correct order suffices.
\end{rmk}

@@

In each of the four cases, when we remove the propagator labelled $p$ we obtain a diagram which satisfies the statement of the theorem by the induction hypothesis, and contains a propagator ${q = (i,j)}$ with no other propagators inside it (with respect to the $<_i$ orientation). Note that this region may or may not contain non-supporting vertices, which we will call $l, l+1, \ldots$ as necessary. Let $V$ be the diagram obtained by removing the propagator $p$ from $W$ (Figure \ref{fig V diagram}). $V$ is guaranteed to be an admissible diagram, since it was formed by removing a propagator from a diagram that was at least weakly admissible.

@@

We first calculate $I^{W}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range (since $p$ will be the final propagator encountered by the algorithm) so if $r \neq p$, then $I_w^W(r) = I_w^V(r)$. Meanwhile, the value of $I^{W}_w(p)$ depends on which step in the algorithm the propagator $q$ contributes a vertex in the diagram $V$, i.e. on the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then 
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority 
%$w \in [l, l+a-1]$} 
%I left this as a comment so we can add it in if necessary, but I don't think we need this level of precision in this sentence: we're just signposting what we're going to do in detail shortly.
of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{W}$, then the remainder of $I_w^{W}$ proceeds identically to the assignments of $I^{V}_{w+1}$ (recall Remark~\ref{rmk algorithm locally same}).  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$}\\
       \end{cases}
       \]
If $j$ is not in the support of $p$, then we must be in Case 2 of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

@@

       m & \text{if $r =p$ and $j$ is not in the support of $p$} \\
       I_{j}^{V}(r) & \text{if $r\neq p$ and $j$ is not in the support of $p$}\end{cases}
       \] 
This completes the analysis of $I_j^W$. Finally, we consider what happens when we start the algorithm at vertex $j+1$. If $j+1$ is in the support of $p$ then we can argue as above to get
\[
       I_{j+1}^{W}(r)  = \begin{cases}
         j+1 & \text{if $r=p$ and $j+1$ is in the support of $p$}\\
         I_{j+2}^{V}(r) & \text{if $r\neq p$ and $j+1$ is in the support of $p$}.
       \end{cases}
       \]
Now suppose $j+1$ is not in the support of $p$. If we start at $j+1$ we need to know whether there are any unassigned propagators supported on edge $i$ when we reach vertex $i$. We already know that $J_q^{V}(j+1) = \{j+1\}$; in particular this means that $q$ contributes $i$ in $I_{j+2}^{V}$, by the induction hypothesis applied to $V$. However the construction of $I^{V}_{j+1}$ first assigns $q$ to $j+1$ and then proceeds identically to $I^{V}_{j+2}$.  In particular if $i$ was assigned in $I^{V}_{j+1}$, then it would not be available to assign to $q$ in $I^{V}_{j+2}$ as all other propagators supported at $i$ in $V$ come before $q$.

>>>>

We summarize here only the portion of positroid theory that we require in this paper; the interested reader is referred to \cite{Postnikov} for more details.

@@

For our purposes, a {\bf positroid} is\footnote{Note that matroid isomorphism allows arbitrary permutations of the ground set but positroid isomorphism allows only cyclic permutations, so the property of being a positroid is not in general preserved by matroid isomorphism.} a matroid $M$ (with ground set $[n]$ and bases set $\cB \subseteq \binom{[n]}{k}$) which can be represented by an element of the positive Grassmannian $\Gr(k,n)$. In other words, there exists a full-rank $k\times n$ real matrix whose $k\times k$ minors are all nonnegative and such that the minor comprised of the columns indicated by the set $J$ 
is positive if and only if $J \in \cB$.

Postnikov shows in \cite{Postnikov} that the positroids of $\Gr(k,n)$ are indexed by many different collections of objects, each with their own advantages and disadvantages. The two most suited to our purposes are Grassmann necklaces and Le diagrams, which we discuss. In order to do so, we first need some preliminary definitions.

@@

\begin{thm} \label{thm WLD defines matroid} \cite[Theorem 3.6]{wilsonloop} The independent sets of the matroid $M(W)$ associated to an admissible Wilson loop diagram $W = (\cP,[n])$ and realized by $C(W)$ are exactly those subsets $V \subseteq [n]$ such that there is no subset $U \subseteq V$ satisfying $|\Prop(U)| < |U|$. \end{thm}
In other words, the independent sets of $M(W)$ correspond to sets of vertices in $W$ in which no subset supports fewer propagators than the vertices it contains.

@@

  Because the nonzero entries of $C(W)$ are independent indeterminants, $J$ is an independent set if and only if there is some choice of $|J|$ nonzero entries of $C(W)$ one in each row associated to an element of $J$ and each in different columns. Each entry in $C(W)$ identifies a propagator by the row of the entry and a vertex by the column of the entry.  The entry is nonzero if and only if the propagator is supported on that vertex.  \hlfix{Consequently, a choice of $|J|$ nonzero entries of $C(W)$, one in each row associated to an element of $J$ and each in different columns, is equivalent to an assignment of the propagators of $J$ to supporting vertices so that no two are assigned to the same vertex.}{Added parenthetical commas according to M.23. better?}  Such an assignment of the propagators of $J$ to supporting vertices is exactly a map $f$ as described in the statement, hence proving the result.
\end{proof}

@@

By relating the behavior of the propagators in $W$ to the matroidal properties of $M(W)$, Agarwala and Marin-Amat also show that $M(W)$ is in fact a positroid whenever $W$ is admissible \cite[Corollary 3.39]{wilsonloop}. \todo{check tenses throughout (M.24)}

However, to go from the Wilson loop diagram $W$ to the positroid cell, one had to go through the matroid explicitly realized via $C(W)$, then make a list of all its nonzero $k\times k$ minors and extract the Le diagram or Grassmann necklace via the methods described in Section~\ref{sec:positroid background}. In \cite{casestudy}, Agarwala and Fryer apply this process in the smallest non-trivial case: admissible Wilson loop diagrams with two propagators on six vertices. Even in this simple case, we see that the mapping from admissible Wilson loop diagrams to positroid cells is neither one-to-one nor onto, and the process described above makes it almost impossible to track the relationship between the original Wilson loop diagram and the resulting positroid.

@@

\begin{dfn}\label{props inside p}
Let $W = (\cP, [n])$ be a weakly admissible Wilson loop diagram, with $p \in \cP$ supported on edges $i$ and $j$. Write $(p, <_i)$ to represent the same propagator directed from $i$ to $j$. Then the sets of propagators {\bf inside} and {\bf outside}  $(p,<_i)$\todo{drop the 'of' throughout. Old: inside of $(p,<_i)$, New: inside $(p,<_i)$} are defined to be
\begin{gather*}\cP_{in}(p,<_i) = \{ (k,l) \in \cP \ |\ i \leq_i k <_i l \leq_i j \}, \\
\cP_{out}(p,<_i) = \cP\setminus \cP_{in}(p,<_i),
\end{gather*}
respectively.
\end{dfn}
Note that we have $p\in \cP_{in}(p, <_i)$ and so $\cP_{in}(p, <_i) = \cP_{out}(p, <_j)\cup\{p\}$.

@@

\begin{lem}\label{lem sian}
  Let $W$ be a weakly admissible Wilson loop diagram with at least 5 vertices and in which each vertex supports at least one propagator.  Then at least one of the following two conditions hold:
  \vspace{-0.5em}
  \begin{enumerate}
    \item $W$ has a propagator of length at most 6 with a propagator of length 2 on one side of it and nothing else on that side.\label{item big and 2} \todo{M.28 may need to be propagated}
    \item There exists a pair of propagators of length $2$ with the property that the first propagator is $(i, i+2)$, the second is $(j, j+2)$, no other propagator has an end between vertices $i+2$ and $j+1$, and $j\in\{i+2, i+3, i+4\}$.\label{item pair of 2s}
  \end{enumerate}
See Figure~\ref{fig: small configs} for an illustration of the two configurations.
\end{lem}

@@

We will inductively construct sequences of propagators $\{p_r\}$ and $\{q_r\}$, with $\ell(p_r) = 2$ for each $r$ and such that either of the following are true:
\begin{itemize}
\item $p_r$ forms part of configuration \ref{item big and 2} or \ref{item pair of 2s} from the statement, at which point the induction terminates, 
  %a configuration \eqref{eq:2 prop config} in $D$
{\em or} 
\item there is a propagator $q_r$ satisfying $\ell(q_r) \geq 4$, and $\{p_1, \ldots, p_r, q_1, \ldots, q_{r-1}\}$ are all on the same side of $q_r$.  
\end{itemize}

@@

The first case is that $i_r \leq j_{r+1} \leq i_r+2$ and $k_r-2 \leq j_{r+1}+2 \leq k_{r}$ (so $p_{r+1}$ has one end before $i_{r} + 3$ and the other after $k_{r} - 2$, i.e. there are at most two vertices between an endpoint of $p_{r+1}$ and an endpoint of $q_r$).  Then since $p_{r+1}$ has length 2, it must be that $(i_r+3)+1 \geq k_r-2$ and so $q_r$ has length $\leq 6$.  By the minimality assumption on $q_r$, no propagator in $W_r$ has $q_r$ on one side and $p_{r+1}$ on the other side. Therefore, any propagator in $W_r$ other than $q_r$ and $p_{r+1}$ must be of the form $(i,j)$ with $i_r\leq i < j\leq i_r+2$ or $k_r-2 \leq i < j\leq k_r$. Both of these require that $(i,j)$ has length 2, and so we have configuration \ref{item pair of 2s} which we have already assumed does not occur.  Consequently, $W_r$ contains only the two propagators $q_r$ and $p_{r+1}$, yielding configuration \ref{item big and 2} from the lemma statement and again contradicting our assumption. Therefore, this first case cannot occur. \todo{M.31 propagate commas after all therefores}

The second case is that there are at least $3$ vertices between one endpoint of $p_{r+1}$ and the nearest endpoint of $q_r$. In other words, at least one of the following four statements are true:
\begin{itemize} 
\item $i_r+3 \leq j_{r+1}\leq k_r-3$ (one endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest endpoint of $q_r$), or
\item $i_r+3 \leq j_{r+1}+2\leq k_r-3$ (the other endpoint of $p_{r+1}$ has at least 3 vertices between it and the nearest endpoint of $q_r$), or
\item $i_r = j_{r+1}$ ($p_{r+1}$ and $q_r$ both have an end on edge $i_r$), or 
\item $k_r = j_{r+1} +2$  ($p_{r+1}$ and $q_r$ both have an end on edge $k_r$).
\end{itemize}
In other words, at least one end of $p_{r+1}$ is supported entirely by vertices in the interval $[i_r+3,k_r+2]$ (the first two bullet points) or both ends lie in $[i_r, i_r+3]$ (third point) or both ends lie in $[k_r-2, k_r+1]$ (fourth point). These situations all behave similarly, and by symmetry it suffices to only consider the second and third possibilities.

We have $j_{r+1}+4 \leq k_r-1$ in both of these cases: in the situation of the second bullet point this follows directly from the inequality, while in the third case it follows from the fact that $\ell(p_{r+1}) = 2$ and $\ell(q_r) \geq 4$. Therefore the vertex $j_{r+1}+4$ is in the interval $[i_r, k_r+1]$ but does not support either $q_r$ or $p_{r+1}$, and must therefore support some other propagator $t$.  Since $p_{r+1}$ is not part of configuration \ref{item pair of 2s} and $W_r$ has no propagators of length 3, it follows that $\ell(t) \geq 4$.  If $q_r$ and $p_{r+1}$ were on different sides of $t$, this would contradict the minimality of $q_r$.  Therefore we can orient $t$ so that all previous $q_i$ and $p_i$ belong to $\cP_{out}(t)$, and set $q_{r+1} = t$ to continue the induction.

The overall result follows by induction. \todo{M.32 find and drop unnecessary 'then's}
\end{proof}

@@

In this section, we give an algorithm for passing directly from a Wilson loop diagram to its Grassmann necklace. This not only greatly simplifies the previously known process to obtain a positroid cell from a Wilson loop diagram, but also allows us to relate the behavior of the positroid $M(W)$ directly to the configuration of propagators in $W$. In Section~\ref{sec poles} we use these results to show that the denominator $R(W)$ of the integrand associated to a Wilson loop diagram can be written in terms of the Grassmann necklace of said diagram.

@@

\begin{dfn} \label{proporderingdfn} \todo{M.36 wants this to be a remark. I disagree? But also note that there's nothing being defined here.}
  There is an ordering on the propagators on $W = (\cP, [n])$ defined by $I_i$. This is the ordering in which they contribute to the algorithm as it constructs $I_i$.
%  $\cP = \{p_1, \ldots p_k\}$, where 
% \[I_i = \{I_i(p_1) <_i I_i(p_2) <_i \ldots <_i I_i(p_k)\}.\]
% (The superscript $W$ is omitted here to simplify the notation.)\note{Sian tweaked this slightly to emphasise the $<_i$ order, because standard convention in most papers is to write all GN terms in the 1-order.}
\end{dfn}

@@

\begin{lem}\label{lem no fourth vertex}
Let $W$ be an admissible Wilson loop diagram containing at least one propagator. For any $i \in [n]$ and for any $p \in \cP$, $I_i(p)$ is not maximal (with respect to $<_i$) amongst the vertices that support $p$.
\end{lem} 
\begin{proof}
Suppose, by way of contradiction that we have $i \in [n]$ and $p \in \cP$ such that $I_i(p)$ is $<_i$-maximal in $V(p)$. There are two possibilities, which are illustrated in Figure~\ref{fig:no fourth vertex}: either $p = (i-1,b)$ for some $b$ (so $i$ and $I_i(p) = i-1$ are the two ends of a single edge), or $p = (a,b)$ with $i\leq_i a <_i b$ and $I_i(p) = b+1$.

When the firt possibility occurs, observe that we must have another propagator $q = (c,b)$ with $c<_i b$ and $I_i(q) = b+1$, because otherwise $I_i$ would assign $p$ to $b+1$. Now $q$ also contributes the $<_i$-maximal vertex in its support to $I_i$ and is an instance of the second possibility. Therefore it suffices to show that no admissible Wilson loop diagram admits the configuration illustrated in Case 2 of Figure~\ref{fig:no fourth vertex}.

@@

In order for $q$ to remain unassigned until the algorithm reaches vertex $b$, there must be another propagator $r$ with an end on edge $d$ and $I_i(r) = d+1$; the only way this can occur is if $r$ is outside of $(q,<_i)$ but inside $(p,<_i)$. But now $r$ contributes its fourth vertex to $I_i$, again contradicting the minimality of $p$.
\end{proof}

@@

\begin{rmk}\label{rmk cyclic}
  Before the proof, let us observe that one of the main ways that the above lemma will be used is in observations like the following.  Suppose $p=(i,j)$ is a propagator and $I_{j+1}(p) = j+1$.  Consider $I_{j+2}(p)$.  Since $j+1$ is the last vertex in the $j+2$ order, $p$ cannot contribute $j+1$ to $I_{j+2}$, so the fact that the intervals in the statement of the lemma are non-empty and occur in their cyclic order implies that $I_{j+2}(p)=i$.  More generally, we know that whenever a propagator stops contributing a particular vertex, its next contribution must be its cyclically next vertex.  This simple observation is surprisingly useful, to the point that the details which appear in the proof of Lemma~\ref{vertex cyclic int lem} are never needed again: the mere fact of non-empty intervals in the correct order suffices.
\end{rmk}

@@

In each of the four cases, when we remove the propagator labelled $p$ we obtain a diagram which satisfies the statement of the theorem by the induction hypothesis, and contains a propagator ${q = (i,j)}$ with no other propagators inside it (with respect to the $<_i$ orientation). Note that this region may or may not contain non-supporting vertices, which we will call $l, l+1, \ldots$ as necessary. Let $V$ be the diagram obtained by removing the propagator $p$ from $W$ (Figure \ref{fig V diagram}). The diagram $V$ is guaranteed to be an admissible diagram, since it was formed by removing a propagator from a diagram that was at least weakly admissible. \todo{Make certain we don't start sentences with simbols}

@@

We first calculate $I^{W}_w$ for a starting vertex ${w \in [n]\backslash \{l, l+1, \ldots,j,j+1\}}$. Note that $p$ has no effect on other propagators for starting vertices in this range (since $p$ will be the final propagator encountered by the algorithm) so if $r \neq p$, then $I_w^W(r) = I_w^V(r)$. Meanwhile, the value of $I^{W}_w(p)$ depends on which step in the algorithm the propagator $q$ contributes a vertex in the diagram $V$, i.e. on the value of $I^{V}_w(q)$. Thus, if $w\in J_q^{V}(i)$ then 
    \[
    I_w^{W}(r) =  \begin{cases}
        \max\{i+1, m\} & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p
      \end{cases} 
    \]
    while if $w\in J_q^{V}(i+1)$ or $w\in J_q^{V}(j)$ then
    \[
    I_w^{W}(r) =  \begin{cases}
        l & \text{if } r=p \\
        I_{w}^{V}(r) & \text{if } r\neq p \;.
      \end{cases} 
      \]
We also need to understand $I^{W}_w$ for $w \in \{l,l+1,\ldots,j,j+1\}$. For the majority 
%$w \in [l, l+a-1]$} 
%I left this as a comment so we can add it in if necessary, but I don't think we need this level of precision in this sentence: we're just signposting what we're going to do in detail shortly.
of these vertices, we use the following observation: if $p$ is the first propagator to be assigned a value by $I_w^{W}$, then the remainder of $I_w^{W}$ proceeds identically to the assignments of $I^{V}_{w+1}$ (recall Remark~\ref{rmk algorithm locally same}).  Thus we have for any $0\leq b <a$
    \[
    I_{l+b}^{W}(r) = \begin{cases}
      l+b & \text{if } r=p\\
      I_{l+b+1}^{V}(r) & \text{if } r\neq p
    \end{cases}
    \]
Similarly, if $j$ is in the support of $p$, then we have
     \[
       I_j^{W}(r) = \begin{cases}
         j & \text{if $r=p$ and $j$ is in the support of $p$}\\
         I_{j+1}^{V}(r) & \text{if $r\neq p$ and $j$ is in the support of $p$} \;.
       \end{cases} 
       \]
If $j$ is not in the support of $p$, then we must be in Case 2 of Figure~\ref{fig 3 cases} with two vertices in the right hand region.  In this case, if we start the algorithm at $j$ we need to know whether there will be any unassigned propagators other than $p$ when we reach vertex $i$, so as to know what $p$ contributes.

@@

       m & \text{if $r =p$ and $j$ is not in the support of $p$} \\
       I_{j}^{V}(r) & \text{if $r\neq p$ and $j$ is not in the support of $p$}\; .\end{cases} 
       \]  \todo{make certain that there are periods after displays that end sentences.}
This completes the analysis of $I_j^W$. Finally, we consider what happens when we start the algorithm at vertex $j+1$. If $j+1$ is in the support of $p$ then we can argue as above to get
\[
       I_{j+1}^{W}(r)  = \begin{cases}
         j+1 & \text{if $r=p$ and $j+1$ is in the support of $p$}\\
         I_{j+2}^{V}(r) & \text{if $r\neq p$ and $j+1$ is in the support of $p$}\;.
       \end{cases}
       \]
Now suppose $j+1$ is not in the support of $p$. If we start at $j+1$ we need to know whether there are any unassigned propagators supported on edge $i$ when we reach vertex $i$. We already know that $J_q^{V}(j+1) = \{j+1\}$; in particular this means that $q$ contributes $i$ in $I_{j+2}^{V}$, by the induction hypothesis applied to $V$. However the construction of $I^{V}_{j+1}$ first assigns $q$ to $j+1$ and then proceeds identically to $I^{V}_{j+2}$.  In particular if $i$ was assigned in $I^{V}_{j+1}$, then it would not be available to assign to $q$ in $I^{V}_{j+2}$ as all other propagators supported at $i$ in $V$ come before $q$.
